---
title: "LDA"
author: "Mike"
date: "2020/5/13"
output: 
  html_document: 
    theme: united
    toc: yes
    toc_float: yes
editor_options: 
  
  
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reference

1. [機器學習: 分類(Classification)-線性區別分析( Linear Discriminant Analysis)](https://medium.com/@chih.sheng.huang821/%E6%A9%9F%E5%99%A8%E5%AD%B8%E7%BF%92-%E9%99%8D%E7%B6%AD-dimension-reduction-%E7%B7%9A%E6%80%A7%E5%8D%80%E5%88%A5%E5%88%86%E6%9E%90-linear-discriminant-analysis-d4c40c4cf937)


1. [Linear Discriminant Analysis -David Valls Lanàquera](https://www.rpubs.com/dvallslanaquera/lda)


1. [Hellinger transformation of vector of abundances](https://stats.stackexchange.com/questions/344760/is-hellinger-transformation-suitable-to-access-species-abundance-environmental)


1. [Ecologically meaningful transformations for ordination of species data](https://link.springer.com/article/10.1007/s004420100716)
---

# Package


```{r message=FALSE, warning=FALSE}
library(dplyr) 
library(magrittr)
library(ggplot2) 
library(vegan) # standardization 
source('Source_LDA.R') # ModelPerformance
```


---

# Data

## Standardization
```{r}
iris.hle <- iris[1:4] %>% as.matrix %>% decostand("hellinger") 
```

## Cluster
```{r}

iris.cluster <- iris.hle %>% vegdist("euc") %>% hclust("ward.D2")
plot(iris.cluster)

iris.gr <-  iris.cluster %>% cutree(3)
table(iris.gr)
```


# Missing values

```{r}
any(is.na(iris))
```

# Multivariate homogeneity

```{r}
iris.pars <- iris[, 1:4] %>% as.matrix
iris.pars.d <- dist(iris.pars)
iris.MHV <- betadisper(iris.pars.d, gr)

iris.MHV
plot(iris.MHV)

anova(iris.MHV)
permutest(iris.MHV)

```

# Boxplot
```{r}
par(mfrow = c(1, ncol(iris.pars)))
for(j in 1:ncol(iris.pars)){
  boxplot(iris.pars[,j], col = (j+1) )
  }
```


# Skewness
```{r}
timeDate::skewness(iris[1:4])
```

# Outliers
```{r}
outliers::outlier(iris[2])
```

```{r}
which(iris[2] >= 4.1 | iris[2] < 2.2) # depend on boxplot
```


# Transformation
```{r}
ptr <- which(iris[2] >= 4.1 | iris[2] < 2.2)
iris[ptr, 2] <- mean(iris[,2])

iris2 <- cbind(log(iris[1]), iris[2], iris[3], iris[4])

```

```{r}
timeDate::skewness(iris2[1:4])

par(mfrow = c(1, ncol(iris2.pars)))
for(j in 1:ncol(iris2.pars)){
  boxplot(iris2.pars[,j], col = (j+1) )
  }


iris2.pars <- iris2[, 1:4] %>% as.matrix
iris2.pars.d <- dist(iris2.pars)
iris2.MHV <- betadisper(iris2.pars.d, gr)

iris2.MHV
par(mfrow = c(1,1))
plot(iris2.MHV)

anova(iris2.MHV)
permutest(iris2.MHV)
```

